install.packages("DBI")
install.packages("RSQLite")
install.packages("dplyr")
install.packages("tidyverse")

#Connecting to the database
```{r}
library(DBI)
library(tidyverse)
library(ggplot2)
library(dplyr)
filename <- "../data/exam_database.db"
my_connection <- dbConnect(RSQLite::SQLite(), dbname = filename)
my_connection
```

#Showing what tables are in the database
```{r}
dbListTables(my_connection)
# We have 3 tables in our database
```

#Reading the database into the global environment
```{r}
con <- dbConnect(RSQLite::SQLite(), "../data/exam_database.db")
film_data <- dbGetQuery(con, "SELECT * FROM films")
```

#Showing what variables (columns) are in each table
```{sql connection=my_connection}
SELECT *
FROM films
LIMIT 10
```
```{sql connection=my_connection}
SELECT *
FROM imdb_scores
LIMIT 10
```
```{sql connection=my_connection}
SELECT *
FROM rotten_tomatoes_scores
LIMIT 10
```

#Confirming the data is between 1970 and 2017 for release_year
```{sql connection=my_connection}
SELECT count(*) AS "Number of Films"
FROM films
JOIN imdb_scores using (film_id)
JOIN rotten_tomatoes_scores using (film_id)
WHERE release_year NOT BETWEEN 1970 AND 2017
ORDER BY audience_score DESC;
```

#Transforming data: mutate genre(s) into genre
```{r}
names(film_data)
names(film_data)[names(film_data) == "genre(s)"] <- "genre"
```

#Finding all genre names 
```{r}
film_data <- dbGetQuery(con, "SELECT * FROM films")
genre_list <- strsplit(film_data$genre, ",")
genre_names <- unique(unlist(genre_list))
print(genre_names)
```

#Finding missing values in rotten tomatoes table
```{sql connection=my_connection}
select 
  sum(film_id is null) as "NULL Film ID",
  sum(audience_score is null) as "NULL Audience Score",
  sum(critic_score is null) as "NULL Critic Score"
from rotten_tomatoes_scores
```

#Finding missing values in imdb table
```{sql connection=my_connection}
select 
  sum(film_id is null) as "NULL Film ID",
  sum(score is null) as "NULL IMDB Score",
  sum(likes is null) as "NULL Likes"
from imdb_scores
```

#Finding missing values in rotten tomatoes table
```{sql connection=my_connection}
select count(*) as "Missing Likes in IMDB"
from imdb_scores
where likes is null;
```

#Counting how many films in imdb scores that has likes
```{sql connection=my_connection}
select count(distinct film_id) as liked_films
from imdb_scores
where "likes" not null
```
#Finding best film with imdb scores
```{sql connection=my_connection}
SELECT films.film_name, AVG(imdb_scores.score) AS avg_imdb_score
FROM imdb_scores
JOIN films ON films.film_id = imdb_scores.film_id
GROUP BY films.film_name
ORDER BY avg_imdb_score DESC
LIMIT 10
```
#Finding best film with audience scores from rotten tomatoes
```{sql connection=my_connection}
SELECT films.film_name, AVG(rotten_tomatoes_scores.audience_score) AS avg_audscore
FROM rotten_tomatoes_scores
JOIN films ON films.film_id = rotten_tomatoes_scores.film_id
GROUP BY films.film_name
ORDER BY avg_audscore DESC
LIMIT 10
```
#Finding best film with average critic scores from rotten tomatoes - Alphabetical Result (NULL)
```{sql connection=my_connection}
SELECT films.film_name, AVG(rotten_tomatoes_scores.critic_score) AS avg_criscore
FROM rotten_tomatoes_scores
JOIN films ON films.film_id = rotten_tomatoes_scores.film_id
GROUP BY films.film_name
ORDER BY avg_criscore DESC
LIMIT 10
```
#Best films based on critic score is maximised and audience score is best
```{sql connection=my_connection} 
SELECT films.*,
       imdb_scores.score AS imdbscores,
       imdb_scores.likes AS imdblikes,
       rotten_tomatoes_scores.audience_score AS rot_audscore,
       rotten_tomatoes_scores.critic_score AS rot_criscore
FROM films
JOIN imdb_scores ON films.film_id = imdb_scores.film_id
JOIN rotten_tomatoes_scores ON films.film_id = rotten_tomatoes_scores.film_id
WHERE critic_score is 100 
ORDER BY audience_score desc
LIMIT 10
```
#Finding worst film with average imdb scores
```{sql connection=my_connection}
SELECT films.film_name, AVG(imdb_scores.score) AS avg_imdb_score
FROM imdb_scores
JOIN films ON films.film_id = imdb_scores.film_id
GROUP BY films.film_name
ORDER BY avg_imdb_score
LIMIT 10
```
#Finding worst film with average audience scores
```{sql connection=my_connection}
SELECT films.film_name, AVG(rotten_tomatoes_scores.audience_score) AS avg_audscore
FROM rotten_tomatoes_scores
JOIN films ON films.film_id = rotten_tomatoes_scores.film_id
GROUP BY films.film_name
ORDER BY avg_audscore 
LIMIT 10
```
#Finding worst film with average critic scores - Alphabetical Result (NULL)
```{sql connection=my_connection}
SELECT films.film_name, AVG(rotten_tomatoes_scores.critic_score) AS avg_criscore
FROM rotten_tomatoes_scores
JOIN films ON films.film_id = rotten_tomatoes_scores.film_id
GROUP BY films.film_name
ORDER BY avg_criscore 
LIMIT 10
```
#Worst films based on critic score is 0 and audience score is worst
```{sql connection=my_connection} 
SELECT films.*,
       imdb_scores.score AS imdbscores,
       imdb_scores.likes AS imdblikes,
       rotten_tomatoes_scores.audience_score AS rot_audscore,
       rotten_tomatoes_scores.critic_score AS rot_criscore
FROM films
JOIN imdb_scores ON films.film_id = imdb_scores.film_id
JOIN rotten_tomatoes_scores ON films.film_id = rotten_tomatoes_scores.film_id
WHERE critic_score is 0 
ORDER BY audience_score 
LIMIT 10
```
#Worst films based on where audience and critics scores are minimal and lowest ranking by audience
```{sql connection=my_connection}
SELECT *
FROM films
JOIN imdb_scores using (film_id)
JOIN rotten_tomatoes_scores using (film_id)
WHERE audience_score BETWEEN 0 AND 65
AND critic_score BETWEEN 0 AND 65
ORDER BY audience_score 
LIMIT 10
```
#Condensed data showing films with Lowest Total scores
```{sql connection=my_connection}
SELECT f.*,
      imdb.score + rt.audience_score + rt.critic_score AS total_score
FROM films AS f
JOIN imdb_scores AS imdb ON f.film_id = imdb.film_id
JOIN rotten_tomatoes_scores AS rt ON f.film_id = rt.film_id
ORDER BY total_score ASC
LIMIT 10
```
#Finding best genre with imdb score
```{sql connection=my_connection}
SELECT films."genre(s)", COUNT(imdb_scores.score) AS imdbcount
FROM films
JOIN imdb_scores ON films.film_id = imdb_scores.film_id
GROUP BY films."genre(s)"
ORDER BY imdbcount desc
LIMIT 10
```
#Finding best genre with average audience score from rotten tomatoes
```{sql connection=my_connection}
SELECT films."genre(s)", AVG(rotten_tomatoes_scores.audience_score) AS count_audscore
FROM rotten_tomatoes_scores
JOIN films ON films.film_id = rotten_tomatoes_scores.film_id
GROUP BY films.film_name
ORDER BY count_audscore desc
```
#Finding best genre with critic score from rotten tomatoes
```{sql connection=my_connection}
SELECT films."genre(s)", AVG(rotten_tomatoes_scores.critic_score) AS count_cri_score
FROM rotten_tomatoes_scores
JOIN films ON films.film_id = rotten_tomatoes_scores.film_id
GROUP BY films.film_name
ORDER BY count_cri_score desc
LIMIT 10
```
#Highest Ranking Genres
```{sql connection=my_connection}
SELECT f."genre(s)" AS "Genre", COUNT(i.score) AS "IMDb Score Count"
FROM films AS f
JOIN imdb_scores AS i ON f.film_id = i.film_id
GROUP BY f."genre(s)"
ORDER BY "IMDb Score Count" DESC
LIMIT 1
```
#Finding worst genre with imdb score
```{sql connection=my_connection}
SELECT films."genre(s)", COUNT(imdb_scores.score) AS imdbcount
FROM films
JOIN imdb_scores ON films.film_id = imdb_scores.film_id
GROUP BY films."genre(s)"
ORDER BY imdbcount 
LIMIT 10
```
#Finding worst genres with average audience score from rotten tomatoes
```{sql connection=my_connection}
SELECT films."genre(s)", AVG(rotten_tomatoes_scores.audience_score) AS count_audscore
FROM rotten_tomatoes_scores
JOIN films ON films.film_id = rotten_tomatoes_scores.film_id
GROUP BY films.film_name
ORDER BY count_audscore
```
#Finding worst genres with critic score from rotten tomatoes
```{sql connection=my_connection}
SELECT films."genre(s)", AVG(rotten_tomatoes_scores.critic_score) AS count_cri_score
FROM rotten_tomatoes_scores
JOIN films ON films.film_id = rotten_tomatoes_scores.film_id
GROUP BY films.film_name
ORDER BY count_cri_score
LIMIT 10
```
#Lowest Ranking Genre
```{sql connection=my_connection}
SELECT f."genre(s)" AS "Genre", COUNT(i.score) AS "IMDb Score Count"
FROM films AS f
JOIN imdb_scores AS i ON f.film_id = i.film_id
GROUP BY f."genre(s)"
ORDER BY "IMDb Score Count"
LIMIT 1
```
#Condensing certain genres
```{sql connection=my_connection}
select "genre(s)",
      case 
           when "genre(s)" in ('Comedy,Romance', 'Romance,Comedy') then 'Rom-Com'
           when "genre(s)" in ('Action,Comedy,Family', 'Family,Comedy,Action') then 'Gritty Family Comedy'
           when "genre(s)" like '%Comedy%' then 'Comedy'
          when "genre(s)" like '%Romance%' then 'Romance'
      end as genres_cleaned
from films as f
join imdb_scores as imdb on f.film_id = imdb.film_id
join rotten_tomatoes_scores as rt on f.film_id = rt.film_id
```
#Plotting Top 5 Genres by Ranking
```{r}
imdb_score_count <- data.frame(
  genre = c("Comedy,Drama,Romance", "Drama", "Comedy", "Comedy,Drama", "Drama,Romance"),
  imdb_score_count = c(244, 216, 199, 198, 158))

ggplot(imdb_score_count, aes(x = reorder(genre, -imdb_score_count), y = imdb_score_count)) +
  geom_bar(stat = "identity", color = "black", fill = "#bada55") +
  xlab("Genre") +
  ylab("IMDB Score Count") +
  ggtitle("Best Genres by Ranking")
```
#Plotting Top 5 Genres by IMDB SCORES Ranking
```{r}
best_film_imdb <- dbGetQuery(con, "
SELECT films.film_name, AVG(imdb_scores.score) AS avg_imdb_score
FROM imdb_scores
JOIN films ON films.film_id = imdb_scores.film_id
GROUP BY films.film_name
ORDER BY avg_imdb_score DESC
LIMIT 5")

ggplot(best_film_imdb, aes(x = film_name, y = avg_imdb_score)) +
  geom_bar(stat = "identity", fill = "green") +
  labs(x = "Film Name", y = "Average IMDB Score") +
  ggtitle("Best Films by IMDB Scores") +
  theme_minimal()
```
#Plotting Worst Genres by Ranking ????
```{r}
imdb_score_count <- data.frame(
  genre = c("Action,Adventure,Fantasy,Action,Adventure,Fantasy", "Action,Biography,Comedy",	"Action,Biography,Western",			"Action,Comedy,Music", "Action,Comedy,Mystery"),
  imdb_score_count = c(1, 1, 1, 1, 1))


ggplot(imdb_score_count, aes(x = reorder(genre, -imdb_score_count), y = imdb_score_count)) +
  geom_bar(stat = "identity", color = "black", fill = "#f55555") +
  xlab("Genre") +
  ylab("IMDB Score Count") +
  ggtitle("Worst Genres by Ranking")
#Note: There are many given scores at 1, precisely from page 1-14 in Lowest Ranking Genres Table
```
#Top 5 Best Films by Total Scores
```{r}
best_film_ts<- dbGetQuery(con, "
SELECT f.*,
       imdb.score as imdb_scores,
       imdb.likes as imdb_likes,
       rt.audience_score as rt_audience_score,
       rt.critic_score as rt_critic_score,
      imdb.score  + rt.audience_score + rt.critic_score as total_score
FROM films as f
JOIN imdb_scores as imdb on f.film_id = imdb.film_id
JOIN rotten_tomatoes_scores as rt on f.film_id = rt.film_id
ORDER BY total_score DESC
LIMIT 5;")

ggplot(best_film_ts, aes(x = film_name, y = total_score)) +
  geom_bar(stat = "identity", fill = "green") +
  labs(x = "Film Name", y = "Total Score") +
  ggtitle("Best Films rated by Total Scores") +
  theme_minimal()
```
#Worst Films by Total Scores
```{r}
worst_filmts_visualisation <- dbGetQuery(con, "
SELECT f.*,
       imdb.score as imdb_scores,
       imdb.likes as imdb_likes,
       rt.audience_score as rt_audience_score,
       rt.critic_score as rt_critic_score,
      imdb.score  + rt.audience_score + rt.critic_score as total_score
FROM films as f
JOIN imdb_scores as imdb on f.film_id = imdb.film_id
JOIN rotten_tomatoes_scores as rt on f.film_id = rt.film_id
ORDER BY total_score
LIMIT 5;")

ggplot(worst_filmts_visualisation, aes(x = film_name, y = total_score)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(x = "Film Name", y = "Total Score") +
  ggtitle("Worst Films rated by Total Scores") +
  theme_minimal()
```
#Lowest 5 Films by IMDB Scores
```{r}
worst_film_imdb <- dbGetQuery(con, "
SELECT films.film_name, AVG(imdb_scores.score) AS avg_imdb_score
FROM imdb_scores
JOIN films ON films.film_id = imdb_scores.film_id
GROUP BY films.film_name
ORDER BY avg_imdb_score
LIMIT 5")

ggplot(worst_film_imdb, aes(x = film_name, y = avg_imdb_score)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(x = "Film Name", y = "Average IMDB Score") +
  ggtitle("Worst Films rated by IMDB Scores") +
  theme_minimal()
```
#Lowest 5 Films by Audience
```{r}
worst_film_aud <- dbGetQuery(con, "
SELECT films.film_name, AVG(rotten_tomatoes_scores.audience_score) AS avg_audscore
FROM rotten_tomatoes_scores
JOIN films ON films.film_id = rotten_tomatoes_scores.film_id
GROUP BY films.film_name
ORDER BY avg_audscore 
LIMIT 5")

ggplot(worst_film_aud, aes(x = film_name, y = avg_audscore)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(x = "Film Name", y = "Average IMDB Score") +
  ggtitle("Worst Films rated by Audience") +
  theme_minimal()
```
